Estimating duration in the presence of misclassification

### 2023-06-07 cis-imperfect-testing.tex
Move test to right section
Outline chapter

In \cref{perf-test}, a framework for analysing the CIS to produce estimates of the duration of positivity was developed.
However, initial application of this framework produced implausibly short estimates; I identified the issue as being due to the presence of false negatives, a form of misclassification bias, within the CIS (\cref{imperf-test:sec:issue}).
There exists little prior work on the incorporation of misclassification bias within similar survival analysis frameworks.
\todo[inline]{Check exactly what has been done and verify exactly what I can claim here}
In this chapter, I modify the previous model to include false negatives (\cref{imperf-test:sec:modelling}).
I show in simulation that this model will recover the duration distribution (\cref{imperf-test:sec:sim-study}) and apply it to the CIS data (\cref{imperf-test:sec:application}).
Plausible estimates of the duration distribution are produced, however, they require assuming that the test sensitivity (the probability of testing positive given being detectable) is known; I discuss what would be required to relax this assumption in \cref{imperf-test:sec:discussion}.

\section{The issues of false negatives} \label{imperf-test:sec:issue}

Demonstrate the likely issue of false negatives (Armitage talk).

Discuss a bit why this is a problem (single positives and bias).

Discuss very low sensitivity needed and changing sensitivity over time (ATACCC and CIS plots).

@@

\subsection{Modifying $p_{ia}$} \label{modifying-p_ia}

@@

\subsection{Modifying $p_{it}$} \label{modifying-p_it}

@@

\section{Simulation study} \label{imperf-test:sec:sim-study}

\subsection{Setup}

>>>>

In \cref{perf-test}, a framework for analysing the CIS to produce estimates of the duration of positivity was developed.
However, initial application of this framework produced implausibly short estimates; I identified the issue as being due to the presence of false negatives, a form of misclassification bias, within the CIS (\cref{imperf-test:sec:problem}).
There exists little prior work on the incorporation of misclassification bias within similar survival analysis frameworks.
\todo[inline]{Check exactly what has been done and verify exactly what I can claim here}
In this chapter, I modify the previous model to include false negatives (\cref{imperf-test:sec:modelling}).
I show in simulation that this model will recover the duration distribution (\cref{imperf-test:sec:sim-study}) and apply it to the CIS data (\cref{imperf-test:sec:application}).
Plausible estimates of the duration distribution are produced, however, they require assuming that the test sensitivity (the probability of testing positive given being detectable) is known; I discuss what would be required to relax this assumption in \cref{imperf-test:sec:discussion}.

\section{The problem} \label{imperf-test:sec:problem}

Applying the method of \cref{perf-test} to CIS data gave implausible estimates of the duration distribution (see \cref{imperf-test:fig:problem-cis-estimates}).
In particular, I estimate a very high probability of the episode lasting less than five days, in conflict with the results of \cref{ATACCC} and other prior work (\eg: \autocite{cevikShedding}).

\begin{figure}
  \centering %\includegraphic{}
  \todo[inline]{Figure showing estimates without false negatives}
  \caption{\label{imperf-test:fig:problem-cis-estimates}}
\end{figure}

Comparing the simulated and CIS data suggested that this is due to the high frequency of single positive episodes in the CIS data.
A single positive episode is one in which the episode contains exactly one positive test, intuitively these cause short estimates because there is no lower bound on the length of time the episode lasts.
Furthermore, due to the truncation, there is little or no ability to upper bound the number of very short episodes because these would almost all be missed.
\Textcite{shenNonparametrica} (building on \textcite{panNote}) showed that in a related setting that the maximum likelihood (frequentist) estimator is inconsistent and can lead to severe underestimation (bias) of the survival function.
While the result is not directly applicable, and can not simply be extended, to the more complex scenario within CIS, it does still build an intuition that these single positive episodes could be causing this issue.
Furthermore, simulations building in even a small number of excess single positive episodes led to problematic estimation even within the CIS setting.
\todo[inline]{My previous application and discussion of \textcite{shenNonparametrica} in this setting would need quite significant reworking to include; is this needed or is this slightly hand-wavy discussion enough?}

Discussion and interrogation of the data led me to believe the most likely reason for an unexpectedly high number of single positive episodes was the presence of false negatives, meaning a false negative as a negative test result returned by a detectable individual.
This can occur for a variety of reasons, most commonly that an individual swabs themselves poorly and hence the concentration of virus on the swab is too low to be detected by the PCR process.
\todo{Maybe I need a better description of the swabbing process?}
Plausibly, a variety of other reasons could also lead to a false negative such as logistical issues or mislabelling of samples.
Consideration of these reasons are beyond the scope of this thesis, and I group all of these possible reasons together.

There are two strands of evidence that support this hypothesis.
First, it is well known that PCR testing can return false negatives \todo{refs for sensitivity} and I needed to explicitly include these in the viral load model of \cref{ATACCC}; they can be directly observed within CIS sa intermittent negatives (see below).
Second, adjusting the simulation to include false negatives can reproduce both the CIS data more realisticly and the same issues when estimating the duration distribution (see \cref{imperf-test:sec:simulate})

\emph{Intermittent negatives} are the clearest example of a false negative.
An intermittent negative is when an individual tests negative but tested positive previously and subsequently.
I explicitly included these in the viral load model of \cref{ATACCC} but ignored in \cref{perf-test}.
Intermittent negatives stripped out when creating the dataset used for all duration analyses in this chapter (as described in \cref{episode-def}), but demonstrate that false negatives do occur.

Whether a future positive is part of the same episode of a reinfection is not always trivial; I rely on a process developed previously by Sarah Walker, see \cref{episode-def}.
However, up until the emergence of the Omicron variant, reinfections are rare, especially in a short time frame, and hence the tricky cases are rare until this time (late 2021).

\section{Simulating false negatives} \label{imperf-test:sec:simulate}

The simplest model for false negatives is by introducing a constant \emph{test sensitivity}, a probability of returning a negative test even when detectable.
This can be simulated by...

However, an implausibly low test sensitivity is required to reproduce the rate of single positive episodes seen in CIS.
\todo[inline]{Figure demonstrating this and references to what reasonable test sensitivities are, including the estimate from the ATACCC chapter}
More plausibly, the rate of false negatives varies over the course of an episode.
False negatives are more likely to occur when the viral load of an individual is low, because there is less virus in their body to be sampled.
Following the short, initial period of growth, viral load largely declines monotonically with time (see \cref{ATACCC}).
\todo[inline]{Reference to ATCCCC chapter, although maybe the specific section with this result would be better}
This suggests that a model with a declining test sensitivity as a function of time since infection might be more suitable.

Examination of the CIS data also provides evidence for a declining test sensitivity.
We can bound the test sensitivity by...
\todo[inline]{Include the analysis of CIS data with bounding test sensitivity}

A simple declining test sensitivty might be:
\todo[inline]{Equation for linearly decreasing test sensitivity}
Simulated data using this test sensitivity is much more similar ot the data, while being more plausible than very low probabilities of false negatives.

\todo[inline]{Add the evidence that false negatives do create the ssame issue in inference in real and simulated data}

@@

\subsection{Modifying \texorpdfstring{$p_{ia}$}{pia}} \label{modifying-p_ia}

@@

\subsection{Modifying \texorpdfstring{$p_{it}$}{pit}} \label{modifying-p_it}

@@

\section{Simulation study results} \label{imperf-test:sec:sim-study-results}
